---
title: "Chapter 1"
author: "Bhupendrasinh Thakre"
date: "3/17/2020"
output: html_document
---

### Chapter 1
## Mean, Median, and Variance

```{r}
library(matrixStats)
state <- read.csv("https://raw.githubusercontent.com/bthakr1/STATS/master/state.csv")

head(state)

# Mean of Population
mean(state[["Population"]])
# Mean of population with trimming on both end by 10%
mean(state[["Population"]], trim = 0.1)
# Median of Population
median(state[["Population"]])
# Weighted Mean of Murder Rate based on Population for respective state
weighted.mean(state[["Murder.Rate"]],w=state[["Population"]])
```

## Variability Estimates

Measures whether the data values are tightly clustered or spared out.

1. Deviation : Difference between observed value and mean estimate of location

2. Variance : Sum of Squared Deviation from the mean divided by n-1, where n is the number of observations

3. Standard Deviation : The square root of variance.

4. Percentile : The value such that P percent of the value take on this value or less and (100-P) 
take on this value or more

Variance, Standard Deviation, and Mean Absolute Deviation are not robust to outliers and extreme values.

5. IQR : Difference between 75th Percentile and 25th Percentile.


```{r}
# Standard Deviation of Population
sd(state[["Population"]])
# Inter Quantile Range of Population
IQR(state[["Population"]])
# Mean Absolute Deviation (Not suspectible to outliers)
mad(state[["Population"]])


```

Spread of Murder Rate based on Quantiles

```{r}
quantile(state[["Murder.Rate"]],p=c(0.05,0.25,0.50,0.75,0.95))

quantile(state[["Murder.Rate"]],p=c(0.10,0.25,0.50,0.75,0.90))
```

Boxplot to visulaize the quantile information

```{r}
boxplot(state[["Population"]]/1000000,ylab="Population in Millions")
```

Frequency Tables

```{r}

breaks <- seq(from=min(state[["Population"]]),to=max(state[["Population"]]), length=11)

pop_freq <- cut(state[["Population"]], breaks = breaks,right = TRUE, include.lowest = TRUE)

breaks <- seq(from=min(state[["Population"]]), to = max(state[["Population"]]), length = 11)
pop_freq <- cut(state[["Population"]], breaks = breaks, right = TRUE, include.lowest = TRUE)
state['PopFreq'] <- pop_freq



```


Histogram : Used to plot bins on X Axis and Count on Y Axis

```{r}

hist(state[["Population"]], breaks = breaks, main = "Histogram of Population", xlab = "Population Bins")

```


Density Plot : Shows the distribution of data values as a continous line.  Sort of "smoothed histogram". 

```{r}
hist(state[["Murder.Rate"]],  xlab = "Murder Rate", ylab = "Frequency", main="Histogram for Murder Rate")
# Density Plot for Murder Rate
hist(state[["Murder.Rate"]], freq = FALSE, , xlab = "Murder Rate", ylab = "Density", main="Density Plot for Murder Rate")
lines(density(state[["Murder.Rate"]]), lwd = 3, col = "blue")

```



# Exploring Binary and Categorical Data

Mode : The most commonly occuring category or value

Bar Chart : The frequency or proportion for each ctegory plotted as bars

``` {r}
# Getting airline Data
library(dplyr)
library(tidyr)
library(ggplot2)
library(corrplot)
library(descr)
airline <- read.csv("https://raw.githubusercontent.com/bthakr1/STATS/master/airline_stats.csv")
dfw <- read.csv("https://raw.githubusercontent.com/bthakr1/STATS/master/dfw_airline.csv")

head(dfw)

barplot(as.matrix(dfw)/6, cex.axis=0.5)


```


Bar Chart vs. Histogram 

In **bar chart** the X axis represents different categories of a factor variable , while in **histogram** X Axis represents value of 
single variable on numeric scale.

**Expected Value : A form of weighted mean. It adds the ideas of future expectations and probability weights , often based on subjective judgements **

# Correlation

Correlation Coefficient : Measures the extent to which numeric variables are associated with one another (-1 to +1)

*Pearson Correlation Coeefieicient* (r): Multiply Deviation of Mean for One Variable with Another and divided by product of Standard 
Deviation

Numerator :

$$\sum_{i=1}^{N}(x_i-\overline{x})(y_i-\overline{y})$$

Denominator :

$$(N-1)s_xs_y$$


Correlation Matrix



```{r}
library(corrplot)
# Based on Pearson Correlation
correlation_mtcars_p <- cor(mtcars, use="complete.obs", method = "pearson")
as.data.frame(correlation_mtcars_p)
corrplot(correlation_mtcars_p, method = "ellipse")
#Based on Spearman Correlation
correlation_mtcars_s <- cor(mtcars, use="complete.obs", method = "spearman")
as.data.frame(correlation_mtcars_s)
corrplot(correlation_mtcars_s, method = "number")

```

Correlation coefficient is sensitive to outliers in the data. 

Other kind of correlations are **Spearman's Rho** and **Kendall's Tau**. These are based on the rank of the data. They are also robust to 
outliers since these worrk on ranks than values. 

Scatte Plot

```{r}

plot(mtcars$mpg, mtcars$wt, xlab = "MPG", ylab = "Weight", main = "Scatter Plot between MPG and Weight")

```

Scatter plots are good when we do not have much data. For larger data sets, we may want to use different way to visuzliae the data.

```{r}
library(ggplot2)
library(hexbin)
kc_tax <- read.csv("https://raw.githubusercontent.com/bthakr1/STATS/master/kc_tax.csv")

kc_tax0 <- subset(kc_tax, kc_tax$TaxAssessedValue < 750000 & kc_tax$SqFtTotLiving > 100 & kc_tax$SqFtTotLiving < 3500)
```

```{r}

# Visualize as hex bins

ggplot(kc_tax0, (aes(x=kc_tax0$SqFtTotLiving,y=kc_tax0$TaxAssessedValue))) + 
  stat_binhex(colour="white") +
  theme_bw() +
  scale_fill_gradient(low = "white", high = "black") +
  labs(x="Finished Square Feet", y ="Tax Assessed Value")
```

To view as Countour Plots

```{r}

ggplot(kc_tax0,(aes(x=kc_tax0$SqFtTotLiving, y = kc_tax0$TaxAssessedValue))) +
  theme_bw() +
  geom_point(alpha = 0.1) +
  geom_density2d(colour="red") +
  labs(x="Finished Square Feet", y ="Tax Assessed Value")


```










































